\documentclass[twoside,openright,12pt]{ociamthesis}  % default square logo 
%\documentclass[12pt,beltcrest]{ociamthesis} % use old belt crest logo
%\documentclass[12pt,shieldcrest]{ociamthesis} % use older shield crest logo

%load any additional packages
\usepackage[top=1in,left=1in,bottom=1in,right=1in]{geometry}
\usepackage{changepage}
\usepackage{lscape}
\usepackage[section]{placeins}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{pgf}
\usepackage{color}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{multirow}
\usepackage{ctable}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage[labelfont={bf,up},font=it]{caption}
%\usepackage{subcaption}
\usepackage{sidecap}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\setcounter{secnumdepth}{0}
\setlength{\parindent}{0cm} 


\newcommand{\T}{\textrm{\textbf{T}}}
\newcommand{\NULL}{\textrm{NULL}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\old}{^\textrm{old}}
\newcommand{\new}{^\textrm{new}}
\newcommand{\vj}{\upsilon_{G_i}}
\newcommand{\st}{Student-\emph{t} }
% \newcommand{\capline1}{\rule{\linewidth}{\arrayrulewidth}}
% \newcommand{\capline2}{\newline\rule{\linewidth}{\arrayrulewidth}}

% For comments in blue.
\newcommand{\blue}[1]{\textsf{\textbf{\textcolor{blue}{[#1]}}}}
% For comments in magenta.
\newcommand{\magenta}[1]{\textsf{\textbf{\textcolor{magenta}{[#1]}}}}
% For comments in red
\newcommand{\red}[1]{\textsf{\textbf{\textcolor{red}{[#1]}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

%% \title{Codimension-Two\\[1ex]     %your thesis title,
%%         Free Boundary Problems}   %note \\[1ex] is a line break in the title

\title{Statistical methods for genotype microarray data  on large cohorts of individuals}
\author{Jared O'Connell}             %your name
\college{Wolfson College}  %your college

%\renewcommand{\submittedtext}{change the default text here if needed}
\degree{the degree of Doctor of Philosophy}     %the degree
\degreedate{Trinity 2013}         %the degree date

%end the preamble and start the document
\supervisor{Supervisor: Dr Jonathan Marchini}
\begin{document}

%this baselineskip gives sufficient line spacing for an examiner to easily
%markup the thesis with comments
\baselineskip=18pt plus1pt

%set the number of sectioning levels that get number and appear in the contents
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}


\maketitle                  % create a title page from the preamble info
%% \include{dedication}        % include a dedication.tex file
%% \include{acknowlegements}   % include an acknowledgements.tex file
%% \include{abstract}          % include the abstract

\begin{abstract}
Genotype microarrays assay hundreds of thousands of genetic variants on an individual's genome. The availability of this high throughput genotyping capability has transformed the field of genetics over the past decade by enabling thousands of individuals to be rapidly assayed.  This has lead to  the discovery of hundreds of genetic variants that are associated with disease and other phenotypes in genome wide association studies (GWAS).  These data have also brought with them a number of new statistical and computational challenges.  This thesis deals with two primary analysis problems involving microarray data; genotype calling and haplotype inference.

Genotype calling involves converting the noisy bivariate fluorescent signals generated by microarray data into genotype values for each genetic variant and individual.  Poor quality genotype calling can lead to false positives and loss of power in GWAS so this is an important task.  We introduce a new genotype calling method that is highly accurate and has the novel capability of fusing microarray data with next-generation sequencing data for greater accuracy and fewer missing values. Our new method compares favourably to other available genotype calling software.

Haplotype inference (or phasing) involves deconvolving these genotypes into the two inherited parental chromosomes for an individual.  The development of phasing methods has been a fertile field for statistical genetics research for well over ten years. Depending on the demography of a cohort, different phasing methods may be more appropriate than others.  We review the popular offerings and introduce a new approach to try and unify two distinct problems; the phasing of extended pedigrees and the phasing of unrelated individuals.  We conduct an extensive   comparison of phasing methods on real and simulated data.  Finally we demonstrate some preliminary results on extending methodology to sample sizes in the tens of thousands.

\end{abstract}

\begin{acknowledgements}
Thank you to my supervisor Dr Jonathan Marchini for his knowledge, guidance and support.

Thanks to SHAPEIT developer, Olivier Delaneau, for many helpful discussions regarding the research in Chapters 3-5.

A shout out to the Genomic Medicine and Statistics 2009 DPhil cohort for all the fun we had together.

Thank you to the Wellcome Trust and the Nuffield Department of
Medicine for the very generous financial support which enabled my
research.

Many thanks to Professor Heather Cordell and Dr Gerton Lunter for
taking the time to examine my thesis and offering helpful comments.
\end{acknowledgements}

\begin{romanpages}          % start roman page numbering
\tableofcontents            % generate and include a table of contents
\listoffigures              % generate and include a list of figures
\listoftables
\end{romanpages}            % end roman page numbering

%now include the files of latex for each of the chapters etc
\doublespacing


\include{introduction}
\include{chapter2}
\include{chapter3}
\include{chapter4}
\include{chapter5}


%next line adds the Bibliography to the contents page
\addcontentsline{toc}{chapter}{Bibliography}
%uncomment next line to change bibliography name to references
%\renewcommand{\bibname}{References}
\bibliography{thesis}
\bibliographystyle{authordate1}

\include{appendix}

\end{document}

